---
title: "data_viz"
author: "Caleb Neale"
date: "4/28/2021"
output: pdf_document
---
# Read Data
```{r}
load(file="data\\cleanedData.Rdata")
```


# Equipment Plotting
## Autoplot variables of interest
```{r}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(fpp3)
library(plotly)

hist(equip$totalHeat)

autoplot(totalEnergyDF, .vars=totalHeat.AHU, ylab="Total AHU Heat, kW")
autoplot(totalEnergyDF, .vars=totalHeat.room, ylab="Total Room Heat, kW")
autoplot(totalEnergyDF, .vars=fanEnergy, ylab="Fan Energy, kW")
autoplot(totalEnergyDF, .vars = totalHeat, ylab="Combined room and AHU Heat, kW")
autoplot(totalEnergyDF, .vars = totalEnergy, ylab="Fan, Room, and AHU Energy, kW") 


# inspect winter break period
totalEnergyDF %>% filter(month(time) == 1 | month(time) == 12) %>% autoplot(.vars = totalEnergy) %>% ggplotly()
totalEnergyDF %>% autoplot(.vars = totalEnergy) %>% ggplotly()


# cut 12-19 to 1-3 as this was a winter break period and doesn't follow the usual cycle
```
## Area Plot Total Energy
```{r}
# get longform data for plotting
totalEnergyDF %>% select(-totalHeat, -totalEnergy) %>% filter(equipment == "AHU2E") %>% pivot_longer(cols = c(totalHeat.AHU, fanEnergy, totalHeat.room), names_to = "datapoint", values_to = "value") -> longAHU2E

totalEnergyDF %>% select(-totalHeat, -totalEnergy) %>% filter(equipment == "AHU2W") %>% pivot_longer(cols = c(totalHeat.AHU, fanEnergy, totalHeat.room), names_to = "datapoint", values_to = "value") -> longAHU2W

# stacked area plots
ggplot(data = longAHU2W, aes(x=time, y=value, fill=datapoint)) + geom_area()
ggplot(data = longAHU2E, aes(x=time, y=value, fill=datapoint)) + geom_area()

# proportional stacked area plots
ggplot(longAHU2W, aes(x=time, y=value, fill=datapoint)) + 
    geom_area(position = "fill") + scale_fill_brewer(palette = "Blues")

ggplot(longAHU2E, aes(x=time, y=value, fill=datapoint)) + 
    geom_area(position = "fill") + scale_fill_brewer(palette = "Blues")
```
## Energy and Temperature
```{r}
weather_merge = weather

# shift weather data by a minute to allow for merging
weather_merge$obsTimeLocal = weather_merge$obsTimeLocal + 60

# merge data
left_join(totalEnergyDF, weather_merge, by=c("time" = "obsTimeLocal")) %>% filter(!is.na(totalEnergy))-> weather_energy

# plot, filtering out low energy values for when HVAC system is "off"
ggplot(data = weather_energy %>% filter(totalEnergy > 50), aes(x=tempAvg, y=totalEnergy)) + geom_point() + geom_smooth()
ggplot(data = weather_energy %>% filter(totalEnergy > 50), aes(x=humidityAvg, y=totalEnergy)) + geom_point() + geom_smooth()

```

## Seasonal plots
```{r}
totalEnergyDF = fill_gaps(totalEnergyDF)
gg_season(totalEnergyDF, totalEnergy, period = "week") %>% ggplotly()
gg_season(totalEnergyDF, totalEnergy, period = "day")
gg_season(totalEnergyDF, totalEnergy, period = "month")
gg_season(totalEnergyDF, totalEnergy, period = "year")
gg_season(totalEnergyDF, totalEnergy, period = "hour")



gg_subseries(fill_gaps(equip), totalHeat, period = "day")

gg_season(fill_gaps(equip), fanEnergy)
gg_season(fill_gaps(equip), fanEnergy, period = "day")
gg_subseries(fill_gaps(equip), fanEnergy, period = "day")

```

# Seasonal Subseries Plots
```{r}
gg_subseries(totalEnergyDF, totalEnergy, period="day")

```

# Lag Plots
```{r}
totalEnergyDF %>% select(totalEnergy) %>% filter(equipment == "AHU2E") %>% gg_lag(y=totalEnergy, geom = "point")
totalEnergyDF %>% select(totalEnergy) %>% filter(equipment == "AHU2W") %>% gg_lag(y=totalEnergy, geom = "point")

```
# Autocorrelation Plots
```{r}

# ACFs
totalEnergyDF %>% filter(equipment == "AHU2E") %>% ACF(totalEnergy) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% ACF(totalEnergy) %>% autoplot()

totalEnergyDF %>% filter(equipment == "AHU2E") %>% ACF(fanEnergy) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% ACF(fanEnergy) %>% autoplot()

totalEnergyDF %>% filter(equipment == "AHU2E") %>% ACF(totalHeat.AHU) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% ACF(totalHeat.AHU) %>% autoplot()

totalEnergyDF %>% filter(equipment == "AHU2E") %>% ACF(totalHeat.room) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% ACF(totalHeat.room) %>% autoplot()

# PACFs
totalEnergyDF %>% filter(equipment == "AHU2E") %>% PACF(totalEnergy) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% PACF(totalEnergy) %>% autoplot()

totalEnergyDF %>% filter(equipment == "AHU2E") %>% PACF(fanEnergy) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% PACF(fanEnergy) %>% autoplot()

totalEnergyDF %>% filter(equipment == "AHU2E") %>% PACF(totalHeat.AHU) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% PACF(totalHeat.AHU) %>% autoplot()

totalEnergyDF %>% filter(equipment == "AHU2E") %>% PACF(totalHeat.room) %>% autoplot()
totalEnergyDF %>% filter(equipment == "AHU2W") %>% PACF(totalHeat.room) %>% autoplot()
```
# Moving average plots
```{r}
totalEnergyMA <- totalEnergyDF %>% mutate(
    `200MA` = slider::slide_dbl(totalEnergy, mean,
                .before =100, .after = 100, .complete = TRUE)
  )

totalEnergyMA %>% filter(equipment == "AHU2E") %>%
  autoplot(totalEnergy, color = "gray") +
  geom_line(aes(y = `200MA`), color = "#D55E00") +
  labs(y = "Total Energy",
       title = "Total Energy over Time, AHU2E") +
  guides(colour = guide_legend(title = "series"))

totalEnergyMA %>% filter(equipment == "AHU2W") %>%
  autoplot(totalEnergy, color = "gray") +
  geom_line(aes(y = `200MA`), color = "#D55E00") +
  labs(y = "Total Energy",
       title = "Total Energy over Time, AHU2W") +
  guides(colour = guide_legend(title = "series"))



```
# Differencing
```{r}
totalEnergyDF %>% filter(equipment == "AHU2E") %>% gg_tsdisplay(totalEnergy, plot_type = "partial")
totalEnergyDF %>% filter(equipment == "AHU2W") %>% gg_tsdisplay(difference(totalEnergy), plot_type = "partial")

```

# Model search
```{r}
models <- totalEnergyDF %>% 
  filter(equipment == "AHU2E") %>% 
  model(stepwise = ARIMA(totalEnergy),
        search = ARIMA(totalEnergy, stepwise = FALSE),
        arima0132 = ARIMA(totalEnergy ~ pdq(0,1,10:32, q_init=32), order_constraint = p + q + P + Q <= 35),
        arima3610 = ARIMA(totalEnergy ~ pdq(10:36,1,0, p_init=36), order_constraint = p + q + P + Q <= 38))


glance(models) %>% arrange(AICc) %>% select(.model:BIC)

models %>% select(search) %>% gg_tsresiduals()

# from book, what are lag and dof parameters? are these values appropriate here?
augment(models) %>%
  filter(.model=='search') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)

# results of above suggest residuals are not white noise, what do I do about this? no next steps listed in the book
```
# Hierarchical approach

## Plotting and formatting
```{r}
# subset and rename energy data

hierEnergy <- totalEnergyDF %>% 
  select(totalHeat.room, fanEnergy, totalHeat.AHU) %>% 
  pivot_longer(names_to = "datapoint", 
               values_to = "value",
               cols = c(totalHeat.room, fanEnergy, totalHeat.AHU)) %>%
  mutate(datapoint = recode(datapoint,
                            `totalHeat.room` = "roomHeat",
                            `totalHeat.AHU` = "AHUHeat",
                            `fanEnergy` = "fanEnergy")
  ) 

as_tsibble(hierEnergy, key = c(datapoint, equipment), index = time) -> temp
                      
# create hierarchical tsibble
hierEnergy %>% aggregate_key(equipment / datapoint, value = sum(value)) -> hierEnergy

# plot hierarchical data
hierEnergy %>%
  autoplot(value) +
  labs(y = "Energy (kW)",
       title = "Energy Consumption, AHU and Components") +
  facet_wrap(vars(datapoint, equipment), scales = "free_y", ncol = 3)

hierEnergy %>%
  autoplot(value) +
  labs(y = "Energy (kW)",
       title = "Energy Consumption, AHU and Components") +
  facet_wrap(vars(equipment), scales = "free_y", ncol = 3)


hierEnergy %>%
  filter(is_aggregated(datapoint)) %>%
  autoplot(value) +
  labs(y = "Energy (kW)",
       title = "Energy Consumption, AHU and Components") +
  facet_wrap(vars(equipment), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")

```
```{r}
# ARIMA
fit <- hierEnergy %>% 
  filter_index(~ "2021-03") %>%
  model(base = ARIMA(value)) %>%
  reconcile(
    bu = bottom_up(base),
    ols = min_trace(base, method = "ols"),
    mint = min_trace(base, method = "mint_shrink"),
  )

fc <- fit %>% forecast(h = "1 month")

fc %>%
  filter(is_aggregated(datapoint)) %>%
  filter_index("2021-04" ~.) %>%
  autoplot(
    hierEnergy %>%
  filter_index("2021-04" ~.),
    level = NULL
  ) +
  labs(y = "Trips ('000)") +
  facet_wrap(vars(equipment), scales = "free_y")

# ETS
fit <- hierEnergy %>% fill_gaps() %>% fill(value, .direction = "downup") %>%
  filter_index(~ "2021-03") %>%
  model(base = ETS(value)) %>%
  reconcile(
    bu = bottom_up(base),
    ols = min_trace(base, method = "ols"),
    mint = min_trace(base, method = "mint_shrink"),
  )

fc <- fit %>% forecast(h = "1 month")

fc %>%
  filter(is_aggregated(datapoint)) %>%
  filter_index("2021-04" ~.) %>%
  autoplot(
    hierEnergy %>%
  filter_index("2021-04" ~.),
    level = NULL
  ) +
  labs(y = "Trips ('000)") +
  facet_wrap(vars(equipment), scales = "free_y")
```

# Regression with ARIMA errors

## Cut winter break data
```{r}
weather_energy$activeHours = 0
weather_energy[weather_energy$totalEnergy > 30, 12] = 1


# cut 12-19 to 1-3 as this was a winter break period and doesn't follow the usual cycle

preWB = weather_energy %>% filter(time < dmy_h("19-12-2020 00"))
postWB =  weather_energy %>% filter(time > dmy_h("3-1-2021 00"))
notWB = rbind(preWB, postWB)

notWB$tempSq = notWB$tempAvg^2
```


## Regression with ARIMA, quadratic on temperature, linear on active hours, not winter break
```{r}
fit1 = notWB %>% 
  filter(equipment == "AHU2E") %>%
  fill_gaps() %>%
  model(ARIMA(totalEnergy ~ tempAvg + tempSq + activeHours))

report(fit1)

bind_rows(
    `Regression residuals` =
        as_tibble(residuals(fit1, type = "regression")),
    `ARIMA residuals` =
        as_tibble(residuals(fit1, type = "innovation")),
    .id = "type"
  ) %>%
  mutate(
    type = factor(type, levels=c(
      "Regression residuals", "ARIMA residuals"))
  ) %>%
  as_tsibble(key=type, index=time) %>%
  gg_season(y=.resid, period = "week")



fit1 %>% gg_tsresiduals()
```

## Include Winter Break Data
```{r}
weather_energy$tempSq = weather_energy$tempAvg^2

fit2 = weather_energy %>% 
  filter(equipment == "AHU2E") %>%
  fill_gaps() %>%
  model(ARIMA(totalEnergy ~ tempAvg + tempSq + activeHours))

report(fit2)

bind_rows(
    `Regression residuals` =
        as_tibble(residuals(fit2, type = "regression")),
    `ARIMA residuals` =
        as_tibble(residuals(fit2, type = "innovation")),
    .id = "type"
  ) %>%
  mutate(
    type = factor(type, levels=c(
      "Regression residuals", "ARIMA residuals"))
  ) %>%
  as_tsibble(key=type, index=time) %>%
  gg_season(y=.resid, period = "week")


fit2 %>% gg_tsresiduals()

```

## Remove nightime data
```{r}
weather_energy %>% filter(totalEnergy > 30) -> temp

fit3 = temp %>% 
  filter(equipment == "AHU2E") %>%
  fill_gaps() %>%
  model(ARIMA(totalEnergy ~ tempAvg + tempSq))

report(fit3)

bind_rows(
    `Regression residuals` =
        as_tibble(residuals(fit2, type = "regression")),
    `ARIMA residuals` =
        as_tibble(residuals(fit2, type = "innovation")),
    .id = "type"
  ) %>%
  mutate(
    type = factor(type, levels=c(
      "Regression residuals", "ARIMA residuals"))
  ) %>%
  as_tsibble(key=type, index=time) %>%
  gg_season(y=.resid, period = "week")


fit3 %>% gg_tsresiduals()

```
## NN Approach
```{r}
fit4 = notWB %>% 
  filter(equipment == "AHU2E") %>%
  fill_gaps() %>%
  fill(totalEnergy, .direction = "downup") %>%
  model(NNETAR(totalEnergy))

report(fit4)

bind_rows(
    `Regression residuals` =
        as_tibble(residuals(fit4, type = "regression")),
    `ARIMA residuals` =
        as_tibble(residuals(fit4, type = "innovation")),
    .id = "type"
  ) %>%
  mutate(
    type = factor(type, levels=c(
      "Regression residuals", "ARIMA residuals"))
  ) %>%
  as_tsibble(key=type, index=time) %>%
  gg_season(y=.resid, period = "week")



fit4 %>% gg_tsresiduals()


fit4 %>%
forecast(h = 48) -> forecasts


fit4 %>%
forecast(h = 200) -> forecasts200



fit4 %>%
  generate(h = 1000, times=100) %>%
  autoplot(.sim) +
  autolayer(notWB %>% filter(time > dmy_h("15-4-2021 00")), totalEnergy) +
  theme(legend.position = "none")


forecasts %>% autoplot(notWB %>% filter(time > dmy_h("15-4-2021 00")))

forecasts200 %>% autoplot(notWB %>% filter(time > dmy_h("15-4-2021 00")))


```

### NN Approach with constrained bounds
```{r}
scaled_logit <- function(x, lower = 0, upper = 300) {
  log((x - lower) / (upper - x))
}
inv_scaled_logit <- function(x, lower = 0, upper = 300) {
  (upper - lower) * exp(x) / (1 + exp(x)) + lower
}
my_scaled_logit <- new_transformation(
                    scaled_logit, inv_scaled_logit)

fit5 <-notWB %>% 
  filter(equipment == "AHU2E") %>%
  fill_gaps() %>%
  fill(totalEnergy, .direction = "downup") %>%
  model(
    NNETAR(
     my_scaled_logit(totalEnergy)
    )
  )


fit5 %>%
  forecast(h=1) -> constrainedForecasts

constrainedForecasts %>%
  autoplot() +
  autolayer(notWB %>% filter(time > dmy_h("15-4-2021 00")), totalEnergy) +
  theme(legend.position = "none")

fit5 %>%
  generate(h = 1000, times=10) %>%
  autoplot(.sim) +
  autolayer(notWB %>% filter(time > dmy_h("15-4-2021 00")), totalEnergy) +
  theme(legend.position = "none")

bind_rows(
    `Regression residuals` =
        as_tibble(residuals(fit5, type = "regression")),
    `ARIMA residuals` =
        as_tibble(residuals(fit5, type = "innovation")),
    .id = "type"
  ) %>%
  mutate(
    type = factor(type, levels=c(
      "Regression residuals", "ARIMA residuals"))
  ) %>%
  as_tsibble(key=type, index=time) %>%
  gg_season(y=.resid, period = "week")


fit5 %>% gg_tsresiduals()

```


# Investigate relationship of building data and enery consupmtion
```{r}
left_join(weather_energy, buildingEnergy, by="time") -> allData

allData$cooling = 0
allData %>% filter(!is.na(tempAvg)) -> allDataTemp
allDataTemp[allDataTemp$tempAvg > 16, 17] = 1

cooling <- allDataTemp %>% filter(cooling == 1) %>% as_tsibble(key = equipment, index = time)
heating <- allDataTemp %>% filter(cooling == 0) %>% as_tsibble(key = equipment, index = time)

ggplot(data = cooling, aes(x=chilledWater, y=totalHeat.AHU)) + geom_point() + geom_smooth(method = "lm")
ggplot(data = heating, aes(x=heating, y= totalEnergy-fanEnergy)) + geom_point() + geom_smooth(method = "lm")

ggplot(data = cooling) + geom_line(color = "red", aes(x=time, y=totalHeat.AHU)) + geom_line(color="blue", aes(x=time, y=(chilledWater*1000)/3412.142))

fig <- plot_ly(cooling, x = ~time, y = ~totalHeat.AHU, name = 'AHU Heat', type = 'scatter', mode = 'lines') 
fig %>% add_trace(y = ~(chilledWater*1000)/3412.142, name = 'Chilled Water', mode = 'lines')

fig <- plot_ly(heating, x = ~time, y = ~ totalEnergy-fanEnergy, name = 'AHU Heat', type = 'scatter', mode = 'lines') 
fig %>% add_trace(y = ~(heating*1000)/3412.142, name = 'Heating', mode = 'lines')

coolingLM = lm(chilledWater~totalHeat.AHU, data = cooling)
heatingLM = lm(heating ~ totalEnergy - fanEnergy, data = heating)

plot(coolingLM$residuals)
plot(heatingLM$residuals)

summary(coolingLM)
summary(heatingLM)

```
```{r}
library(mgcv)
notWB$Day = day(notWB$time)
gamm(totalEnergy ~ s(tempAvg) + s(tempSq) + activeHours + s(time), data=notWB, correlation = corARMA(form = ~ 1|Day, p = 1))
```











```{r}
weather_energy %>% select(totalEnergy) %>% filter(equipment == "AHU2E") -> temp2

temp2 = temp2[,1:2]



ts(temp2, start = 1595602800, frequency = 48) -> timeseries

plot(timeseries)


models <- model(stepwise = ARIMA(timeseries),
        search = ARIMA(timeseries, stepwise = FALSE),
        arimaLarge = ARIMA(timeseries ~ pdq(1:50,1,1:50, q_init=32, p_init=36), order_constraint = p + q + P + Q <= 110))

search = ARIMA(timeseries, stepwise = FALSE)
arimaLarge = ARIMA(timeseries ~ pdq(1:50,1,1:50, q_init=32, p_init=36), order_constraint = p + q + P + Q <= 110)

library(forecast)

auto.arima(timeseries[,1]) %>% forecast(h=40) %>% plot()



  autoplot() +
  autolayer(notWB %>% filter(time > dmy_h("15-4-2021 00")), totalEnergy) +
  theme(legend.position = "none")

search %>% gg_tsresiduals()

```

