---
title: "Data Cleaning and (preliminary) EDA"
author: "Caleb Neale"
date: "Spring 2021"
output:
  pdf_document: default
  html_notebook: default
subtitle: Optimizing HVAC Operation for Occupant Comfort and Energy Savings
editor_options: 
  chunk_output_type: inline
---
# HVAC System Summary
System diagram + data accessible through:
http://icoweb.fm.virginia.edu/anyglass/pubdisplay/UVa/Customers/LinkLab0202/Home.gdfx

Acronyms decoded using below resources:

 - https://www.abraxasenergy.com/energy-resources/toolbox/hvac-acronyms/
 - https://github.com/uva-eng-time-series-sp21/neale-caleb/blob/main/HVAC_System_Documentation/141602%20-%20M0.0.pdf


### For each air handling unit we have:

- Temperature data (deg F):
 - RA-T; Return Air
 - SA-T; Supply Air
 - PH-T; Pre-Heat Air
 - MA-T; Mixed Air

- Fans (logical on/off, Variable Frequency Drive perecntage):
  - R-FN; Return Fan
  - S-FN; Supply Fan

- Humidity:
  - RA-H; Return Air (% rel. humidity)

- Ducts:
  - EXH-D; Exhaust (%, open? documentation unclear)
  - OA-D; Outside Air (%, open? documentation unclear)

- Heating/Cooling:
  - PH-V; Pre-Heat Valve (% Open)
  - CHW-V; Chilled Water Valve (% Open)

- Occupied (logical 0/1)

### For each room we have:
- HW-V; Hot Water Valve (%, open? documentation unclear)
- SA-T; Supply Air Temperature (73.3 deg F)
- SA-F; Supply Air Flow (CFM)
- SA-F-SP; Supply Air Flow Set-Point
- ZN-T; Unconfirmed but appears to be temperature set-point (deg F)
- Temperature (deg C)
- co2 (only select rooms, PPM)


# Load libraries 
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(fpp3)
```

# Import Data and convert to tibble
```{r warning=FALSE}
read_and_clean <- function(csv_path){
  df <- read.csv(csv_path, sep=";", row.names = NULL)
  colnames(df) <- c("series", 'time','value')
  
  # NAs will be induced by following line, seems like this occurs when the value in the 'value' column == null
  df$value <- as.numeric(df$value)
  df<- df[-1,]
  df <- as_tibble(df)
  return(df)
}

co2 <- read_and_clean('co2.csv')
occupied_status <- read_and_clean('occupied_status.csv')
occupied_status$value <- as.factor(occupied_status$value)
supply_air_flow <- read_and_clean('supply_air_flow.csv')
supply_fan <- read_and_clean('supply_fan.csv')
supply_fan$value <- as.factor(supply_fan$value)
temperature <- read_and_clean('temperature.csv')

```
# Check for NaN
```{r}
sum(is.na(occupied_status$value))
sum(is.na(co2$value))
sum(is.na(supply_air_flow$value)) # lot of NAs (over 32000)
sum(is.na(supply_fan$value))
sum(is.na(temperature$value)) # lot of NAs (over 15000)

```

# Convert time data to datetime format
```{r}

convert_to_datetime <- function(df){
  df$time <- gsub("-04:00$", "-0400", df$time)
  df$time <- gsub("-05:00$", "-0500", df$time)
  df$time <- strptime(df$time, format ="%Y-%m-%dT%H:%M:%S%z")
  df$time <- as.POSIXct(df$time)
  return(df)
}

co2 <- convert_to_datetime(co2)
occupied_status <- convert_to_datetime(occupied_status)
supply_air_flow <- convert_to_datetime(supply_air_flow)
supply_fan <- convert_to_datetime(supply_fan)
temperature <- convert_to_datetime(temperature)

```


# Investigate data in series columns
```{r}
co2 %>% count(series) 

supply_air_flow %>% count(series)

supply_fan %>% count(series)

temperature %>% count(series)

occupied_status %>% count(series)

```
Co2 data is only provided for 6 rooms: Olsson 203, 211, 213, 217, 221, 225. 

Supply_air_flow contains data for 45 rooms in Olsson hall, as well as set-point data for each room. (Note: Investigate documentation for definition of set-point data)

Temperature is given for 44 rooms in Olsson, all but the generic "2nd floor" label which was found in the supply_air_flow table. There is no set-point data provided here. 

Supply_fan_status is given for both HVAC units; the nature of the time intervals of the supply generating process is still under investigation. 

Occupied_status is given for both HVAC units. As the status is not given by room, I'm looking for documentation which shows what occupied_status means in the system. The nature of the time intervals of the supply generating process is still under investigation. 


# Create rooms column by parsing from series column 
```{r}
co2$room = regmatches(x= co2$series, m=regexpr("([0-9]{3})", co2$series))

supply_air_flow$room = str_match(supply_air_flow$series, "C[0-9]{3}|[0-9]{3}")

temperature$room = str_match(temperature$series, "C[0-9]{3}|[0-9]{3}")
```

There exists a mapping from HVAC unit to rooms in the plans for the HVAC system (HVAC_System_Documentation folder in file "141605 - M0.3.pdf", tables on right side) which could be used to relate observations on the room level and observations on the system level (e.g. which rooms are receiving supply at a given time based on supply_status data).


# Create room assignment vectors for each HVAC unit

```{r}
AHU_2E <- c(241, 243, 245, 247, 249, 251, 253, 257, 255, 259, 263, 261, 240, "C244", 244, 260, 213, 217, 225, 218, "C230", 220, "C210", "T212", "T218", "C216", "C214", "T210", 256, 229, 231, 223, "C227", "C211", 254, "C250", 258) 

AHU_2W <- c(269, 267, 265, 273, 271, 275, 277, 279, 281, 283, 285, 274, 286, 204, 208, 272, 270, "C260", "C200", "C201", 203, 276, "C280", "C270", 211, 201)

rooms_tbl <- rbind(tibble('room' = AHU_2E, 'equipment' = "AHU_2E"), tibble('room' = AHU_2W, 'equipment' = "AHU_2W"))

# Check for duplicates/overlap
rooms_tbl %>% group_by(room) %>% count() %>% filter(n>1) -> dupes 

print(dupes)
```
No duplicates.

# Parse equipment names from series column
```{r}
occupied_status$equiment <- str_match(occupied_status$series, "AHU2[EW]")
supply_fan$equipment     <- str_match(supply_fan$series, "AHU2[EW]")
```


# Table Designs


## Table for room-specific indoor environmental quality data

Index: date-time, 3 hour intervals

Keys: room

Observations: 
 
 * CO$_2$ levels (ppm, mean)
 * temperature (degrees C)
 * supply air flow (ft^3 s-1), 
 * supply air flow set-point (ft^3 s-1), 

Data cleaning tasks:

- Unstack supply_air_flow data such that there is a column for value and a column for set-point for each room at each time-stamp
- Aggregate flow and set-point values for each HVAC unit to create an HVAC unit value at each time-stamp
- For every three hour interval, assign the supply fan status column to the most recent value from supply_fan for each AHU
- Calculate energy consumption and input into final column 

### Unstack supply_air_flow

```{r}
supply_air_flow %>% filter(grepl("Setpoint", series)) -> supply_air_flow_setpoints

supply_air_flow %>% filter(!grepl("Setpoint", series)) -> supply_air_flow
```


```{r join ieq data to unified table}
co2 %>%
  rename('co2_ppm_mean' = 'value') %>%
  select(time, room, co2_ppm_mean) -> co2_ts

temperature %>%
  rename('temperature_C' = 'value') %>%
  select(time, room, temperature_C) -> temperature_ts

supply_air_flow %>%
  rename('supply_air_flow_cfs' = 'value') %>%
  select(time, room, supply_air_flow_cfs) -> supply_air_flow_ts

supply_air_flow_setpoints %>%
  rename('supply_air_flow_setpoint_cfs' = 'value') %>%
  select(time, room, supply_air_flow_setpoint_cfs) -> supply_air_flow_setpoints_ts

full_join(co2_ts, temperature_ts, by=c("time", "room")) %>%
  full_join(supply_air_flow_ts, by=c("time", "room")) %>%
  full_join(supply_air_flow_setpoints_ts, by=c("time", "room")) -> ieq_tbl

print(ieq_tbl)
```

## Table of equipment operational data

Index: date-time, irregular time intervals

Keys: HVAC equipment code

Observations: 

 * supply fan (logical)
 * occupancy status (logical)
 * aggregated air flow (ft^3/min)
 
```{r}
supply_fan %>%
  rename('supply_fan_status' = 'value') %>%
  select(time, equipment, supply_fan_status) -> supply_fan_irreg_ts

occupied_status %>%
  rename('occupied' = 'value',
         'equipment' = 'equiment') %>%
  select(time, equipment, occupied) -> occupied_irreg_ts

equipment_ops_irreg_ts <- full_join(supply_fan_irreg_ts, occupied_irreg_ts, by=c("time", "equipment")) 

print(equipment_ops_irreg_ts %>% arrange(time))

```


Q: Do rooms that share HVAC equipment also share the same supply air flow and setpoints?

```{r do rooms that share HVAC equipment also share same supply air flow and setpoints}

ieq_tbl %>%
  filter(room %in% (rooms_tbl %>% filter(equipment == "AHU_2E") %>% pull(room))) %>%
  select(time, room, supply_air_flow_cfs, supply_air_flow_setpoint_cfs) %>%
  arrange(time, room)

```
A: Clearly not. Rooms can share the same HVAC equipment yet have different setpoints and different realized supply air flows.


```{r}
inner_join(supply_air_flow_setpoints,supply_air_flow, by=c("room", "time")) %>% select(c("time", "value.x", "value.y", "room")) -> unstacked_supply_air

colnames(unstacked_supply_air) <- c("time", "setpoint", "value", "room")

```

** Q: How does energy consumption related to reported values of system operations -- esp. supply air flow, supply fan status, occupied status? **

#### Formulae

SEER = BTU/Watt-Hours = total change in heat energy in conditioned space divided by energy consumed to condition space


##### For heating:

BTU/hr = C * del(T) * M

Where:
- C is the specific heat of air (approx. 1 kJ/Kg*K)
- del(T) is the change in temperature
- M is the mass of air = rho*V, using an approximate density of 1.2 Kg/m^3
- Given our data an adjustment will also have to be made to convert from cubic feet per **minute** to BTU per **hour**.


##### For cooling, an additional term can be included in calculating BTU related to the removal of humidity:

BTU/hr = C * del(T) * M + (0.68 * CFM * del(w_gr))

Note about above: This concept makes sense but I have been unable to find sourcing outside of a blog post on an HVAC website. 

- CFM = cubic feet per minute
- del(w_gr) = change in humidiy ratio in grains

I have been unable to locate a data source on the differential between supply and return air humidity so I will be excluding this term for now, with the potential for revisiting later as there is one measure of relative humidity as a part of return air that is available to glean information from. 


#### Data

Data to be exported from icoweb platform:

- Temperature differential (del(T)) can be calculated by finding the difference between relevant points in the AHU, which I believe to be supply and return air (RA-T minus SA-T)


### Aggregate flow and set point by HVAC unit

```{r}
unstacked_supply_air <- as_tsibble(unstacked_supply_air, key= room, index = time)

unstacked_supply_air %>% filter(room %in% AHU_2E) -> air_supply_AHU_2E
unstacked_supply_air %>% filter(room %in% AHU_2W) -> air_supply_AHU_2W

aggregated_AHU_2E <- aggregate(cbind(air_supply_AHU_2E$setpoint, air_supply_AHU_2E$value), by=list(time=air_supply_AHU_2E$time), FUN=sum)

aggregated_AHU_2W <- aggregate(cbind(air_supply_AHU_2W$setpoint, air_supply_AHU_2W$value), by=list(time=air_supply_AHU_2W$time), FUN=sum)

colnames(aggregated_AHU_2E) = c("time", "setpoint", "air_supply")
colnames(aggregated_AHU_2W) = c("time", "setpoint", "air_supply")

inner_join(aggregated_AHU_2E, aggregated_AHU_2W, by=c("time"), suffix=c(".AHU2E", ".AHU2W")) -> unit_supply_air

```

### Calculate estimated energy usage
```{r}

```


### Feed forward supply_fan data
I did finally get this to work, but further research into the factors which contribute to energy consumption suggest that supply fan status is not a significant contributor to energy consumption.
```{r, eval=FALSE}
# create df where columns are time, AHU2E status, AHU2W status
supply_fan %>% filter(equipment == "AHU2E") -> AHU2E_supply_fan
supply_fan %>% filter(equipment == "AHU2W") -> AHU2W_supply_fan

unstacked_fan = full_join(AHU2E_supply_fan, AHU2W_supply_fan, by=c("time"))
unstacked_fan = as.data.frame(unstacked_fan)

unstacked_fan %>% select(2,3,6) -> unstacked_fan
colnames(unstacked_fan) = c("time", "AHU2E_status", "AHU2W_status")
unstacked_fan$time = as.POSIXlt(unstacked_fan$time)

# down fill dataset 
unstacked_fan %>% fill(AHU2E_status, .direction ="down") -> unstacked_fan
unstacked_fan %>% fill(AHU2W_status, .direction ="down") -> unstacked_fan


time_difference = function(time, times){
  if(!is.na(time)){
    differences = as.data.frame(times - time)
    colnames(differences) = c("value")
    pos_differences = dplyr::filter(differences, value > 0)$value
    pos_diff_index = which(as.numeric(pos_differences) == min(as.numeric(pos_differences)))
    rel_time = times[which(differences$value == pos_differences[pos_diff_index])]

    unit_supply_air[unit_supply_air$time == rel_time,"AHU2W_fan_status"] <<- unstacked_fan[unstacked_fan$time==time, "AHU2W_status"]
    unit_supply_air[unit_supply_air$time == rel_time,"AHU2E_fan_status"] <<- unstacked_fan[unstacked_fan$time==time, "AHU2E_status"]
  }
  else{
    print("fail")
  }
}

times = unit_supply_air$time


for (i in 1:length(unstacked_fan$time)) {
  time = unstacked_fan$time[i]
  time_difference(time, times)
}

# down fill dataset 
unit_supply_air %>% fill(AHU2E_fan_status, .direction ="down") -> unit_supply_air
unit_supply_air %>% fill(AHU2W_fan_status, .direction ="down") -> unit_supply_air

```


## Table for analysis of system dynamics and room comfort
This analysis is being set aside for a later time

Key: room, Index: time

Observations: 

* c02, 
* occupied status, 
* supply air flow, 
* supply fan, 
* temperature, 
* supply air flow set-point



# Convert to tsibble objects
```{r}
co2 <- as_tsibble(co2, key= series, index = time)
occupied_status <- as_tsibble(occupied_status, key= series, index = time)
supply_air_flow <- as_tsibble(supply_air_flow, key= room, index = time)
supply_fan <- as_tsibble(supply_fan, key= series, index = time)
temperature <- as_tsibble(temperature, key= series, index = time)

```

