---
title: "Data Cleaning and (preliminary) EDA"
author: "Caleb Neale"
date: "Spring 2021"
output:
  pdf_document: default
  html_notebook: default
subtitle: Optimizing HVAC Operation for Occupant Comfort and Energy Savings
---

# Load libraries 
```{r}
library(tidyverse)
library(lubridate)
library(fpp3)
```

# Import Data and convert to tibble
```{r}
read_and_clean <- function(csv_path){
  df <- read.csv(csv_path, sep=";", row.names = NULL)
  colnames(df) <- c("series", 'time','value')
  df$value <- as.numeric(df$value)
  df<- df[-1,]
  df <- as_tibble(df)
  return(df)
}

co2 <- read_and_clean('co2.csv')
occupied_status <- read_and_clean('occupied_status.csv')
supply_air_flow <- read_and_clean('supply_air_flow.csv')
supply_fan <- read_and_clean('supply_fan.csv')
temperature <- read_and_clean('temperature.csv')

```

# Convert time data to datetime format
```{r}

convert_to_datetime <- function(df){
  df$time <- gsub("-04:00$", "-0400", df$time)
  df$time <- gsub("-05:00$", "-0500", df$time)
  df$time <- strptime(df$time, format ="%Y-%m-%dT%H:%M:%S%z")
  df$time <- as.POSIXct(df$time)
  return(df)
}

co2 <- convert_to_datetime(co2)
occupied_status <- convert_to_datetime(occupied_status)
supply_air_flow <- convert_to_datetime(supply_air_flow)
supply_fan <- convert_to_datetime(supply_fan)
temperature <- convert_to_datetime(temperature)

```

# Convert to tsibble objects
```{r}
co2 <- as_tsibble(co2, key= series, index = time)
occupied_status <- as_tsibble(occupied_status, key= series, index = time)
supply_air_flow <- as_tsibble(supply_air_flow, key= series, index = time)
supply_fan <- as_tsibble(supply_fan, key= series, index = time)
temperature <- as_tsibble(temperature, key= series, index = time)

```

# Investigate data in series columns
```{r}
co2 %>% count(series)

supply_air_flow %>% count(series)

supply_fan %>% count(series)

temperature %>% count(series)

occupied_status %>% count(series)

```
Co2 data is only provided for 6 rooms: Olsson 203, 211, 213, 217, 221, 225. Investigating whether additional rooms are available. 

Supply_air_flow contains data for 45 rooms in Olsson hall, as well as set-point data for each room. (Note: Investigate documentation for definition of set-point data)

Supply_fan_status is given for both HVAC units; the nature of the time intervals of the supply generating process is still under investigation. 

Temperature is given for 44 rooms in Olsson, all but the generic "2nd floor" label which was fond in the supply_air_flow table. There is no set-point data provided here. 

Occupied_status is given for both HVAC units. As the status is not given by room, I'm looking for documentation which shows what occupied_status means in the system. The nature of the time intervals of the supply generating process is still under investigation. 


# Create rooms column by parsing from series column 
```{r}
co2$room = regmatches(x= co2$series, m=regexpr("([0-9]{3})", co2$series))

supply_air_flow$room = str_match(supply_air_flow$series, "C[0-9]{3}|[0-9]{3}")

temperature$room = str_match(temperature$series, "C[0-9]{3}|[0-9]{3}")

```

There exists a mapping from HVAC unit to rooms which could be used to relate observations on the room level and observations on the system level (e.g. which rooms are receiving supply at a given time based on supply_status data)

# Create room assignment vectors for each HVAC unit

```{r}
AHU_2E <- c(241, 243, 245, 247, 249, 251, 253, 257, 255, 259, 263, 261, 240, "C244", 244, 260, 213, 217, 225, 218, "C230", 220, "C210", "T212", "T218", "C216", "C214", "T210", 256, 229, 231, 223, 225, "C227", "C211", 254, "C250", 258) 

AHU_2W <- c(269, 267, 265, 273, 271, 275, 277, 279, 281, 283, 285, 274, 286, 204, 208, 272, 270, "C260", "C200", "C201", 203, 276, "C280", "C270", 211, 201)

# Check for duplicates/overlap

table(AHU_2E)

table(AHU_2W)

table(c(AHU_2E, AHU_2W))

```
# Final Table Designs

## Table for analysis of system dynamics and energy consumption

Key: HVAC unit, index: time (3 hour intervals? pending documentation)
Observations: supply air flow (aggregated by HVAC unit), supply air flow set-point (aggregated by HVAC unit), supply fan, energy use (calculated)


- Unstack supply_air_flow data such that there is a column for value and a column for set-point for each room at each time-stamp
- Aggregate flow and set-point values for each HVAC unit to create an HVAC unit value at each time-stamp
- For every three hour interval, assign the supply fan status column to the most recent value from supply_fan
- Calculate energy consumption and input into final column 

### Unstack supply_air_flow

```{r}
supply_air_flow %>% filter(grepl("Setpoint", series)) -> supply_air_flow_setpoints

supply_air_flow %>% filter(!grepl("Setpoint", series)) -> supply_air_flow


inner_join(supply_air_flow_setpoints,supply_air_flow, by=c("room", "time")) %>% select(c("time", "value.x", "value.y", "room")) -> unstacked_supply_air

colnames(unstacked_supply_air) <- c("time", "setpoint", "value", "room")

```



## Table for analysis of system dynamics and room comfort

Key: room, Index: time
Observations: c02, occupied status, supply air flow, supply fan, temperature, supply air flow set-point



## Remaining tasks

- ?? From every row of supply_fan, create a row for each room which maps to that HVAC unit with the correct supply_status
- Obtain more detailed documentation on occupied_status to determine if above approach would be appropriate, apply as necessary
- Determine binning strategy for 
- Merge tsibbles to create final tsibble
_ 

