---
title: "Data Cleaning and (preliminary) EDA"
author: "Caleb Neale"
date: "Spring 2021"
output:
  pdf_document: default
  html_notebook: default
subtitle: Optimizing HVAC Operation for Occupant Comfort and Energy Savings
---

# Load libraries 
```{r}
library(tidyverse)
library(lubridate)
library(fpp3)
```

# Import Data and convert to tibble
```{r}
read_and_clean <- function(csv_path){
  df <- read.csv(csv_path, sep=";", row.names = NULL)
  colnames(df) <- c("series", 'time','value')
  df$value <- as.numeric(df$value)
  df<- df[-1,]
  df <- as_tibble(df)
  return(df)
}

co2 <- read_and_clean('co2.csv')
occupied_status <- read_and_clean('occupied_status.csv')
supply_air_flow <- read_and_clean('supply_air_flow.csv')
supply_fan <- read_and_clean('supply_fan.csv')
temperature <- read_and_clean('temperature.csv')

```

# Convert time data to datetime format
```{r}

convert_to_datetime <- function(df){
  df$time <- gsub("-04:00$", "-0400", df$time)
  df$time <- gsub("-05:00$", "-0500", df$time)
  df$time <- strptime(df$time, format ="%Y-%m-%dT%H:%M:%S%z")
  df$time <- as.POSIXct(df$time)
  return(df)
}

co2 <- convert_to_datetime(co2)
occupied_status <- convert_to_datetime(occupied_status)
supply_air_flow <- convert_to_datetime(supply_air_flow)
supply_fan <- convert_to_datetime(supply_fan)
temperature <- convert_to_datetime(temperature)

```

# Convert to tsibble objects
```{r}
co2 <- as_tsibble(co2, key= series, index = time)
occupied_status <- as_tsibble(occupied_status, key= series, index = time)
supply_air_flow <- as_tsibble(supply_air_flow, key= series, index = time)
supply_fan <- as_tsibble(supply_fan, key= series, index = time)
temperature <- as_tsibble(temperature, key= series, index = time)

```

# Investigate data in series columns
```{r}
co2 %>% count(series)

supply_air_flow %>% count(series)

supply_fan %>% count(series)

temperature %>% count(series)

occupied_status %>% count(series)

```
Co2 data is only provided for 6 rooms: Olsson 203, 211, 213, 217, 221, 225. Investiating whether additional rooms are available. 

Supply_air_flow contains data for 45 rooms in Olsson hall, as well as setpoint data for each room. (Note: Investigate documentation for definition of setpoint data)

Supply_fan_status is given for both HVAC units; the nature of the time intervals of the supply generating process is still under investigation. 

Temperature is given for 44 rooms in Olsson, all but the generic "2nd floor" label which was fond in the supply_air_flow table. There is no setpoint data provided here. 

Occupied_status is given for both HVAC units. As the status is not given by room, I'm looking for documentation which shows what occupied_status means in the system. The nature of the time intervals of the supply generating process is still under investigation. 


# Create rooms column by parsing from series column 
```{r}
co2$room = regmatches(x= co2$series, m=regexpr("([0-9]{3}) Olsson", co2$series))

supply_air_flow$room = str_match(supply_air_flow$series, "[(](.+)[)]")[,2]

temperature$room = str_match(temperature$series, "(([0-9]{3}) Olsson)|(C([0-9]{3}) Olsson)")

```

There exists a mapping from HVAC unit to rooms which could be used to relate observations on the room level and observations on the system level (e.g. which rooms are recievin supply at a given time based on supply_status data)

# Final Table Design

Key: room, Index: time
Observations: c02, occupied status, supply air flow, supply fan, temperature, supply air flow setpoint

## Remaining tasks

- ?? From every row of supply_fan, create a row for each room which maps to that HVAC unit with the correct supply_status
- Obtain more detailed documentation on occupied_status to determine if above approach would be appropriate, apply as necessary
- Determine binning strategy for 
- Merge tsibbles to create final tsibble
_ 

